Manuscript Figures

### 2021-09-17 src/manuscript_figures.Rmd
add tumor purity estimation comparison plot and update figure legend

p2 <- ggplot(stats[purity_method == 'CPE'][grep('TCGA', project)], 
       aes(x = reorder(project, COR), y = COR, fill = COR)) + 
  geom_bar(stat = 'identity') +
  coord_flip() +
  theme(axis.title.y = element_blank(),  legend.position = 'none') + 
  labs(y = "MAUI vs CPE") + ylim(0, 1)

# correlation plots for top cancer types
selected <- stats[purity_method == 'IHC'][grep('TCGA', experiment)][order(COR, decreasing = T)][1:4]$experiment
plots <- lapply(selected, function(x) {
  message(x)
  p <- cowplot::plot_grid(plotlist = lapply(names(results[[x]]), function(y) {
    message(y)
    ggpubr::ggscatter(results[[x]][[y]][['preds']], 
                           x = 'y_hat', y = 'y', 
                           add = 'reg.line', cor.coef = T, cor.method = 'pearson') + 
      labs(x = paste('MAUI prediction of', y), y = y, title = gsub("\\..+$", "", x))
  }))
  return(p)
})

p3 <- cowplot::plot_grid(plotlist = plots, ncol = 1)

# supplementary figure
supFig <- cowplot::plot_grid(cowplot::plot_grid(p1, p2, labels = c('A', 'B')), 
                             p3, nrow = 2, labels = c('', 'C'), 
                             rel_heights = c(2, 5))

# print p3 to supplements 
supFig_count <- supFig_count + 1
SuppFigures[[supFig_count]] <- supFig
fig.name <- paste0('SuppFigure', supFig_count, '.tumor_purity.supp.pdf')
names(SuppFigures)[supFig_count] <- fig.name
ggsave(file.path(FigureFolder, fig.name), 
              plot = supFig, units = 'in', width = 8.27, height = 11.69)
SuppLegends[[supFig_count]] <- "Predicting tumor purity estimates using MAUI latent factors. 
A) Correlation of the MAUI predictions of IHC versus the annotated IHC values (immunohistochemistry-based tumor purity estimation). 
B) Correlation of the MAUI predictions of CPE versus the annotated CPE levels (CPE stands for consensus purity estimated by computational methods: ABSOLUTE, ESTIMATE, and LUMP). 
C) Scatterplots of MAUI predictions versus the annotated IHC and CPE values for the 
top best predicted cancer types."
```

>>>>

# compute other tools' correlations with IHC
tp <- get_tumor_purity_estimates()
absolute <- tp[!is.na(ABSOLUTE), mean(cor(ABSOLUTE, IHC, use = 'complete.obs')),by = project]
estimate <- tp[!is.na(ESTIMATE), mean(cor(ESTIMATE, IHC, use = 'complete.obs')),by = project]
lump <- tp[!is.na(LUMP), mean(cor(LUMP, IHC, use = 'complete.obs')),by = project]

dt <- stats[purity_method == 'IHC']
colnames(dt)[3] <- 'MAUI'
dt$ABSOLUTE <- absolute[match(dt$project, project)]$V1
dt$LUMP <- lump[match(dt$project, project)]$V1
dt$ESTIMATE <- estimate[match(dt$project, project)]$V1

mdt <- melt(dt, measure.vars = c('MAUI', 'ABSOLUTE', 'ESTIMATE', 'LUMP'), id.vars = 'project')
p2 <- ggboxplot(mdt, x = 'variable', y = 'value', add = 'jitter', color = 'variable') +
  theme(legend.position = 'none', axis.title.x = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(y = 'Average Correlation with IHC')

p3 <- ggplot(stats[purity_method == 'CPE'][grep('TCGA', project)], 
       aes(x = reorder(project, COR), y = COR, fill = COR)) + 
  geom_bar(stat = 'identity') +
  coord_flip() +
  theme(axis.title.y = element_blank(),  legend.position = 'none') + 
  labs(y = "MAUI vs CPE") + ylim(0, 1)

# correlation plots for top cancer types
selected <- stats[purity_method == 'IHC'][grep('TCGA', experiment)][order(COR, decreasing = T)][1:2]$experiment
plots <- lapply(selected, function(x) {
  message(x)
  p <- cowplot::plot_grid(plotlist = lapply(names(results[[x]]), function(y) {
    message(y)
    ggpubr::ggscatter(results[[x]][[y]][['preds']], 
                           x = 'y_hat', y = 'y', 
                           add = 'reg.line', cor.coef = T, cor.method = 'pearson') + 
      labs(x = paste('MAUI prediction of', y), y = y, title = gsub("\\..+$", "", x))
  }))
  return(p)
})

p4 <- cowplot::plot_grid(plotlist = plots, ncol = 1)

# supplementary figure
supFig <- cowplot::plot_grid(cowplot::plot_grid(p1, p2, p3, labels = c('A', 'B', 'C'), nrow = 1), 
                             p4, nrow = 2, labels = c('', 'D'))

# print p3 to supplements 
supFig_count <- supFig_count + 1
SuppFigures[[supFig_count]] <- supFig
fig.name <- paste0('SuppFigure', supFig_count, '.tumor_purity.supp.pdf')
names(SuppFigures)[supFig_count] <- fig.name
ggsave(file.path(FigureFolder, fig.name), 
              plot = supFig, units = 'in', width = 8.27, height = 11.69)
SuppLegends[[supFig_count]] <- "Predicting tumor purity estimates using MAUI latent factors. 
A) Correlation of the MAUI predictions of IHC versus the annotated IHC values (immunohistochemistry-based tumor purity estimation). 
B) Comparison of average correlations between computational estimates of IHC values by different tools. Each point represents the 
value obtained for each of the TCGA cohorts seen in panel A. ABSOLUTE tool lacks estimates for 6 out of 14 cancer types. 
C) Correlation of the MAUI predictions of CPE versus the annotated CPE levels (CPE stands for consensus purity estimated by computational methods: ABSOLUTE, ESTIMATE, and LUMP). 
D) Scatterplots of MAUI predictions versus the annotated IHC and CPE values for the 
top two best predicted cancer types with respect to annotated IHC values."
```
